---
title: "Ggplot practice"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(here)
library(janitor)
library(ggplot2)
library(paletteer)
```

Ggplot basics

```{r}

lizards <- read_csv(here("data_tidy", "lizards.csv"))
```

## the essentials

-   data

-   aesthetics

-   geometry

```{r}

# ggplot essentials, 3 ways
ggplot(data = lizards,
       aes(x = total_length, y = weight)) + geom_point()


# Second 
ggplot(data = lizards) +
  geom_point(aes(x = total_length, y = weight))

# Third
ggplot() +
  geom_point(data = lizards,
             aes(x = total_length, y = weight))
```

## Swap out geometries

```{r}
# bad example
ggplot(data = lizards,
       aes(x = total_length, y = weight)) + geom_line()
```

Some geometry types only need axis

```{r}
ggplot(data = lizards, aes(x = total_length)) + geom_histogram()

```

Jitterplots usually are best with one categorical and one continuous variable

```{r}
ggplot(data = lizards, aes(y = common_name, x = weight)) +
  geom_jitter()
```

## Aesthetic mapping

### Updating based on a constant value

don't need to use the aes() function

```{r}
ggplot(data = lizards, aes(x = weight)) +
  geom_histogram(color = "orange",
                 fill = "purple",
                 alpha = 0.3,
                 linetype = "dotted")
```

## Updating aesthetics based on variables

In that case, we need to use the aes() function

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name, size = total_length),
             alpha = 0.3)

```

# Themes

## Update axis labels

```{r}
ggplot(data = lizards, aes(x = weight, y = site)) +
  geom_jitter(aes(color = common_name)) +
  theme_minimal() +
  labs(x = "Weight (g)",
       y = "Study site",
       title = "Lizard weights across sites at Jornada", color = "species") # Trocar título da legenda
```

## Reorder data

```{r}
lizards_count <- lizards %>% 
  count(common_name)

ggplot(data = lizards_count,
       aes(x = n, y = fct_reorder(common_name, n))) + 
  geom_col() +
labs(x = "Individuals (n)",
     y = "Species common name")
```

## Advanced ggplot customization

```{r}
ggplot(data = lizards,
       aes(x = total_length, y = weight)) +
  geom_point() +
  scale_x_continuous(breaks = c(0, 250, 500), 
                     limits = c(0, 500)) +
  scale_y_continuous(breaks = seq(from = 0,
                                  to = 70,
                                  by = 10),
                     limits = c(0, 70)) +
  labs(x = "Total length (mm)",
       y = "Weight (grams)") +
  theme_minimal()


```

```{r}
# Counting up the number of lizards caught per day
lizards_count <- lizards %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  count(date)

ggplot(data = lizards_count, aes(x = date, y = n)) +
  geom_line() +
  scale_x_date(breaks = scales::breaks_width("3 years"), # See date_breaks for next year!
               labels = scales::label_date("'%y")) + # See date_labels for next year!
  scale_y_log10(labels = scales::label_scientific())
```

### Customization of aesthetics with scale\_ functiond

```{r}
# color gradient with just 2 colors
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) +
  scale_color_gradient(low = "red", high = "navy")


```

```{r}
# color gradient with more than 2 colors
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) +
  scale_color_gradientn(colors = c("orange", "red", "purple", "navy", "black"))
```

```{r}
# Create a gradient of color bins
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) +
  scale_color_steps(low = "red", high = "navy", breaks = seq(0,80, by = 5))
```

```{r}
ggplot(data = lizards,
       aes(x = common_name, y = total_length)) +
  geom_boxplot(aes(fill = common_name),
               show.legend = F) +
  coord_flip()

```

```{r}
# ordering species based on median length
lizards_mean <- lizards %>% 
  mutate(common_name = fct_reorder(common_name, total_length, .fun = median))

# Then make a graph (use View(palettes_d_names) to see other discrete palettes in {paletteer})

ggplot(data = lizards_mean, aes(y = common_name, x = total_length)) +
  geom_boxplot(aes(fill = common_name), show.legend = FALSE) +
  scale_x_continuous(limits = c(0, 500)) +
  scale_fill_paletteer_d("lisa::AlbrechtDurer") +
  labs(y = "Lizard species",
       x = "Total length (mm)") +
  theme_minimal() +
  theme(panel.grid.minor = element_blank())
```

```{r}
library(patchwork)

p1 <- ggplot(data = lizards_mean, aes(y = common_name, x = total_length)) +
  geom_boxplot(aes(fill = common_name), show.legend = FALSE) +
  scale_x_continuous(limits = c(0, 500)) +
  scale_fill_paletteer_d("lisa::AlbrechtDurer") +
  labs(y = "Lizard species",
       x = "Total length (mm)") +
  theme_minimal() +
  theme(panel.grid.minor = element_blank())

```

```{r}
p2 <- ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = weight)) +
  scale_color_steps(low = "red", high = "navy", breaks = seq(0,80, by = 5))
```

```{r}
library(patchwork)

# "|" puts plots next to each other
(p1 | p2) & theme_minimal()
# "/" puts plots on top of each other
(p1 / p2) & theme_minimal()

# Combination of next to and on top of each other
p3 <- ((p1 | p2) / p2) & theme_minimal()

ggsave("test.png", p3)
```

```{r}

```
