---
title: "Ggplot practice"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(here)
library(janitor)
```

Ggplot basics

```{r}

lizards <- read_csv(here("data_tidy", "lizards.csv"))
```

## the essentials

-   data

-   aesthetics

-   geometry

```{r}

# ggplot essentials, 3 ways
ggplot(data = lizards,
       aes(x = total_length, y = weight)) + geom_point()


# Second 
ggplot(data = lizards) +
  geom_point(aes(x = total_length, y = weight))

# Third
ggplot() +
  geom_point(data = lizards,
             aes(x = total_length, y = weight))
```

## Swap out geometries

```{r}
# bad example
ggplot(data = lizards,
       aes(x = total_length, y = weight)) + geom_line()
```

Some geometry types only need axis

```{r}
ggplot(data = lizards, aes(x = total_length)) + geom_histogram()

```

Jitterplots usually are best with one categorical and one continuous variable

```{r}
ggplot(data = lizards, aes(y = common_name, x = weight)) +
  geom_jitter()
```

## Aesthetic mapping

### Updating based on a constant value

don't need to use the aes() function

```{r}
ggplot(data = lizards, aes(x = weight)) +
  geom_histogram(color = "orange",
                 fill = "purple",
                 alpha = 0.3,
                 linetype = "dotted")
```

## Updating aesthetics based on variables

In that case, we need to use the aes() function

```{r}
ggplot(data = lizards, aes(x = total_length, y = weight)) +
  geom_point(aes(color = common_name, size = total_length),
             alpha = 0.3)

```

# Themes

## Update axis labels

```{r}
ggplot(data = lizards, aes(x = weight, y = site)) +
  geom_jitter(aes(color = common_name)) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))
  labs(x = "Weight (g)",
       y = "Study site",
       title = "Lizard weights across sites at Jornada", color = "species") # Trocar título da legenda
```

## Reorder data

```{r}
lizards_count <- lizards %>% 
  count(common_name)

ggplot(data = lizards_count,
       aes(x = n, y = fct_reorder(common_name, n))) + 
  geom_col()
labs(x = "Individuals (n)",
     y = "Species common name")
```
